---
title: "biol792ch2"
author: "morgan hrivnak"
date: "2024-09-03"
output: word_document
---

2e1. (2) pr(rain|monday)

2e2. (3) the probability that it is monday, given that it is raining

2e3. (1) pr(monday|rain)

2e4. If you were to spin a globe and randomly point to any part of the globe their is a 70% chance that you would be pointing to an area of water.

2m1. 
```{r}
# WWW
# define grid
p_grid <- seq( from=0 , to=1 , length.out=100 )

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=3 , prob=p_grid )

# define prior (uniform)
prior <- rep(1,100)
posterior <- likelihood * prior
postierior <- posterior / sum(posterior)

plot( posterior ~ p_grid , type="l" )

```
```{r}
# WWWL
# define grid
p_grid <- seq( from=0 , to=1 , length.out=100 )

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=4 , prob=p_grid )

# define prior (uniform)
prior <- rep(1,100)
posterior <- likelihood * prior
postierior <- posterior / sum(posterior)

plot( posterior ~ p_grid , type="l" )

```
```{r}
# LWWLWWW
# define grid
p_grid <- seq( from=0 , to=1 , length.out=100 )

# compute likelihood at each value in grid
likelihood <- dbinom( 5 , size=7 , prob=p_grid )

# define prior (uniform)
prior <- rep(1,100)
posterior <- likelihood * prior
postierior <- posterior / sum(posterior)

plot( posterior ~ p_grid , type="l" )
```
```{r}
# WWWL
# define grid
p_grid <- seq( from=0 , to=1 , length.out=100 )

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=3 , prob=p_grid )

# define prior (uniform)
prior <- rep(1,100)
posterior <- likelihood * prior
postierior <- posterior / sum(posterior)

plot( posterior ~ p_grid , type="l" )
``
```

2m2.

```{r}
#WWW
# define grid
p_grid <- seq( from=0 , to=1 , length.out=100 )

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=3 , prob=p_grid )

# define new prior 
prior <- ifelse( p_grid < 0.5 , 0 , 1 )
posterior <- likelihood * prior
postierior <- posterior / sum(posterior)

plot( posterior ~ p_grid , type="l" )
```
```{r}
#WWWL
# define grid
p_grid <- seq( from=0 , to=1 , length.out=100 )

# compute likelihood at each value in grid
likelihood <- dbinom( 3 , size=4 , prob=p_grid )

# define new prior 
prior <- ifelse( p_grid < 0.5 , 0 , 1 )
posterior <- likelihood * prior
postierior <- posterior / sum(posterior)

plot( posterior ~ p_grid , type="l" )
```
```{r}
#LWWLWWW
# define grid
p_grid <- seq( from=0 , to=1 , length.out=100 )

# compute likelihood at each value in grid
likelihood <- dbinom( 5 , size=7 , prob=p_grid )

# define new prior 
prior <- ifelse( p_grid < 0.5 , 0 , 1 )
posterior <- likelihood * prior
postierior <- posterior / sum(posterior)

plot( posterior ~ p_grid , type="l" )
```

2m3. P(E|L)=P(L|E)⋅P(E)P(L)
            =P(L|E)⋅P(E)P(L|E)⋅P(E)+P(L|M)⋅P(M)
            =0.3⋅0.50.3⋅0.5+1.0⋅0.5
            ≈0.2308
            
2m4. P(b=B|f=B)=P(f=B,b=B)P(f=B)
               =(1/3) / (3/6)
               =2/3

2m5. P(b = B|f = B)= (1/2)/(5/8) = 4/5

2m6. 

card|# of cards|# of black sides|# ways to|frequency
b/b   1               2               2       50%
b/w   2               1               2       50%
w/w   3               0               0       0

2m7. 

C1 | C2  | # of ways to see (B,W) | frequency
b/b| b/w           2*1                 2/8
b/b| w/w           2*2                 4/8
b/w| b/b           1*0                 0
b/w| w/w           1*2                 2/8   
w/w| b/b           0*0                 0   
w/w| b/w           0*1                 0

