---
title: "biol7392ch14hw"
author: "morgan hrivnak"
date: "2024-11-21"
output:
  html_document:
    df_print: paged
---
14E1

The original B now needs to be a Bgroup because we want the slope to vary not just the intercept group. With this we need to then define agroup and Bgroup as a joint distribution. B is now the average slope like how a is the average intercept. These both neede to be defined as agroupBgroup ~ MVNormal([aB],S). S defines how they are linked. Like how we need to define priors we also need to define covariance matrix as S= (sigmaa 00 sigmaB) R (sigmaa 00 sigmaB)

14E2. 

You could have a situation that causes high intercepts and steep slopes. Biologically speaking we can think about having a larger population of birds reproducing at faster rates as they have more and more offspring. If carrying capacity didn't exist or was no where in site for a species they would have more and more offspring. This situation is not very likely to happen biologically. 

14E3. 

When there is not a lot of variation it can cause shrinkage in cluster variation. Adaptive regularization can cause the data to shrink even more because their are too many parameters and they are then not effectively been utilized.

14M1. 
```{r}
library(rethinking)
library(tidyverse)
library(rstan)
library(dplyr)
library(tidybayes)

a <- 3.5 # average morning wait time
b <- (-1) # average difference afternoon wait time
sigma_a <- 1 # std dev in intercepts
sigma_b <- 0.5 # std dev in slopes
rho <- 0 # correlation between intercepts and slopes
Mu <- c(a, b)
cov_ab <- sigma_a * sigma_b * rho
Sigma <- matrix(c(sigma_a^2, cov_ab, cov_ab, sigma_b^2), ncol = 2)
# simulate observations
N_cafes <- 20
set.seed(6)
vary_effects <- mvrnorm(N_cafes, Mu, Sigma)
a_cafe <- vary_effects[, 1]
b_cafe <- vary_effects[, 2]
N_visits <- 10
afternoon <- rep(0:1, N_visits * N_cafes / 2)
cafe_id <- rep(1:N_cafes, each = N_visits)
mu <- a_cafe[cafe_id] + b_cafe[cafe_id] * afternoon
sigma <- 0.5 # std dev within cafes
wait <- rnorm(N_visits * N_cafes, mu, sigma)
# package into  data frame
d <- data.frame(cafe = cafe_id, afternoon = afternoon, wait = wait)

m_M1 <- ulam(
  alist(
    wait ~ normal(mu, sigma),
    mu <- a_cafe[cafe] + b_cafe[cafe] * afternoon,
    c(a_cafe, b_cafe)[cafe] ~ multi_normal(c(a, b), Rho, sigma_cafe),
    a ~ normal(5, 2),
    b ~ normal(-1, 0.5),
    sigma_cafe ~ exponential(1),
    sigma ~ exponential(1),
    Rho ~ lkj_corr(2)
  ),
  data = d, chains = 4, cores = 4
)

post <- extract.samples(m_M1)
ggplot() +
  stat_halfeye(aes(x = post$Rho[, 1, 2])) +
  theme_bw() +
  labs(x = "Rho")

```

14M2.
```{r}
# set up parameters of population
a <- 3.5 # average morning wait time
b <- -1 # average difference afternoon wait time
sigma_a <- 1 # std dev in intercepts
sigma_b <- 0.5 # std dev in slopes
rho <- -0.7 # correlation between intercepts and slopes
Mu <- c(a, b)
cov_ab <- sigma_a * sigma_b * rho
Sigma <- matrix(c(sigma_a^2, cov_ab, cov_ab, sigma_b^2), ncol = 2)
# simulate observations
N_cafes <- 20
set.seed(42)
vary_effects <- mvrnorm(N_cafes, Mu, Sigma)
a_cafe <- vary_effects[, 1]
b_cafe <- vary_effects[, 2]
N_visits <- 10
afternoon <- rep(0:1, N_visits * N_cafes / 2)
cafe_id <- rep(1:N_cafes, each = N_visits)
mu <- a_cafe[cafe_id] + b_cafe[cafe_id] * afternoon
sigma <- 0.5 # std dev within cafes
wait <- rnorm(N_visits * N_cafes, mu, sigma)
# package into  data frame
d <- data.frame(cafe = cafe_id, afternoon = afternoon, wait = wait)

m_M2Baseline <- ulam(
  alist(
    wait ~ normal(mu, sigma),
    mu <- a_cafe[cafe] + b_cafe[cafe] * afternoon,
    c(a_cafe, b_cafe)[cafe] ~ multi_normal(c(a, b), Rho, sigma_cafe),
    a ~ normal(5, 2),
    b ~ normal(-1, 0.5),
    sigma_cafe ~ exponential(1),
    sigma ~ exponential(1),
    Rho ~ lkj_corr(2)
  ),
  data = d, chains = 4, cores = 4)

m_M2 <- ulam(
  alist(
    wait ~ dnorm(mu, sigma),
    mu <- a_cafe[cafe] + b_cafe[cafe] * afternoon,
    a_cafe[cafe] ~ dnorm(a, sigma_alpha),
    b_cafe[cafe] ~ dnorm(b, sigma_beta),
    a ~ dnorm(0, 10),
    b ~ dnorm(0, 10),
    sigma ~ dexp(1),
    sigma_alpha ~ dexp(1),
    sigma_beta ~ dexp(1)
  ),
  data = d, chains = 4, cores = 4)

post_Base <- extract.samples(m_M2Baseline)
a_Base <- apply(post_Base$a_cafe, 2, mean)
b_Base <- apply(post_Base$b_cafe, 2, mean)
post_M2 <- extract.samples(m_M2)
a_M2 <- apply(post_M2$a_cafe, 2, mean)
b_M2 <- apply(post_M2$b_cafe, 2, mean)
plot(a_M2, b_M2,
  xlab = "intercept", ylab = "slope",
  pch = 16, col = rangi2, ylim = c(min(b_M2) - 0.05, max(b_M2) + 0.05),
  xlim = c(min(a_M2) - 0.1, max(a_M2) + 0.1), cex = 2
)

```
The model that knows about the correlation is better because it accounts for more information about the entire population and mirrors exactly what was generated at the same time.

14M3. 
```{r}
data(UCBadmit)
d <- UCBadmit
dat_list <- list(
  admit = d$admit,
  applications = d$applications,
  male = ifelse(d$applicant.gender == "male", 1, 0),
  dept_id = rep(1:6, each = 2)
)
str(dat_list)

Begin_C <- Sys.time()
m_M3 <- ulam(
  alist(
    admit ~ dbinom(applications, p),
    logit(p) <- a[dept_id] + bm[dept_id] * male,
    c(a, bm)[dept_id] ~ multi_normal(c(a_bar, bm_bar), Rho, sigma_dept),
    a_bar ~ dnorm(0, 1),
    bm_bar ~ dnorm(0, 1),
    sigma_dept ~ dexp(1),
    Rho ~ dlkjcorr(2)
  ),
  data = dat_list, chains = 4, cores = 4)

End_C <- Sys.time()
precis(m_M3, 3)

Begin_NC <- Sys.time()
m_M3NonCent <- ulam(
  alist(
    admit ~ dbinom(applications, p),
    logit(p) <- a_bar + v[dept_id, 1] + (bm_bar + v[dept_id, 2]) * male,
    transpars > matrix[dept_id, 2]:v <- compose_noncentered(sigma_dept, L_Rho, z),
    matrix[2, dept_id]:z ~ dnorm(0, 1),
    a_bar ~ dnorm(0, 1.5),
    bm_bar ~ dnorm(0, 1),
    vector[2]:sigma_dept ~ dexp(1),
    cholesky_factor_corr[2]:L_Rho ~ lkj_corr_cholesky(2)
  ),
  data = dat_list, chains = 4, cores = 4)

End_NC <- Sys.time()
precis(m_M3NonCent, 3)

#centered data
End_C - Begin_C

#non centered data
End_NC - Begin_NC

```
The centered model is more efficient.
The non-centered model ran faster.

12M4.
```{r}
data(Kline)
d <- Kline
# Chapter 11 stuff
d$P <- scale(log(d$population))
d$contact_id <- ifelse(d$contact == "high", 2, 1)
# Chapter 14 stuff
d$society <- 1:10
data(islandsDistMatrix)

## Chapter 11 Model
dat2 <- list(T = d$total_tools, P = d$population, cid = d$contact_id)
m11.11 <- ulam(
  alist(
    T ~ dpois(lambda),
    lambda <- exp(a[cid]) * P^b[cid] / g,
    a[cid] ~ dnorm(1, 1),
    b[cid] ~ dexp(1),
    g ~ dexp(1)
  ),
  data = dat2, chains = 4, cores = 4, log_lik = TRUE)

# Chapter 14 Model
dat_list <- list(T = d$total_tools, P = d$population, society = d$society, Dmat = islandsDistMatrix)
m14.8 <- ulam(
  alist(
    T ~ dpois(lambda),
    lambda <- (a * P^b / g) * exp(k[society]),
    vector[10]:k ~ multi_normal(0, SIGMA),
    matrix[10, 10]:SIGMA <- cov_GPL2(Dmat, etasq, rhosq, 0.01), c(a, b, g) ~ dexp(1), etasq ~ dexp(2), rhosq ~ dexp(0.5)
  ),
  data = dat_list, chains = 4, cores = 4, iter = 2000, log_lik = TRUE)

compare(m11.11, m14.8)

```

14M5.
```{r}
data(Primates301)
d <- Primates301
d$name <- as.character(d$name)
dstan <- d[complete.cases(d$group_size, d$body, d$brain), ]
spp_obs <- dstan$name
dat_list <- list(
  N_spp = nrow(dstan),
  M = standardize(log(dstan$body)),
  B = standardize(log(dstan$brain)),
  G = standardize(log(dstan$group_size)),
  Imat = diag(nrow(dstan))
)

data(Primates301_nex)
tree_trimmed <- keep.tip(Primates301_nex, spp_obs) # only keep tree that's relevant to our species
Rbm <- corBrownian(phy = tree_trimmed) # calculate expected covariance given a Brownian model
V <- vcv(Rbm) # compute expected variances and covariances
Dmat <- cophenetic(tree_trimmed) # cophenetic distance matrix
dat_list$V <- V[spp_obs, spp_obs] # covariances in speciesXspecies matrix
dat_list$R <- dat_list$V / max(V) # relative covariances of speciesXspecies matrix

m_M5Ordi <- ulam(
  alist(
    G ~ multi_normal(mu, SIGMA),
    mu <- a + bM * M + bB * B,
    matrix[N_spp, N_spp]:SIGMA <- Imat * sigma_sq,
    a ~ normal(0, 1),
    c(bM, bB) ~ normal(0, 0.5),
    sigma_sq ~ exponential(1)
  ),
  data = dat_list, chains = 4, cores = 4)

m_M5Brown <- ulam(
  alist(
    G ~ multi_normal(mu, SIGMA),
    mu <- a + bM * M + bB * B,
    matrix[N_spp, N_spp]:SIGMA <- R * sigma_sq,
    a ~ normal(0, 1),
    c(bM, bB) ~ normal(0, 0.5),
    sigma_sq ~ exponential(1)
  ),
  data = dat_list, chains = 4, cores = 4)

dat_list$Dmat <- Dmat[spp_obs, spp_obs] / max(Dmat)
m_M5GP <- ulam(
  alist(
    G ~ multi_normal(mu, SIGMA),
    mu <- a + bM * M + bB * B,
    matrix[N_spp, N_spp]:SIGMA <- cov_GPL1(Dmat, etasq, rhosq, 0.01),
    a ~ normal(0, 1),
    c(bM, bB) ~ normal(0, 0.5),
    etasq ~ half_normal(1, 0.25),
    rhosq ~ half_normal(3, 0.25)
  ),
  data = dat_list, chains = 4, cores = 4)

plot(coeftab(m_M5Ordi, m_M5Brown, m_M5GP), pars = c("bM", "bB"))


```
When you do not take phylogency into account there is a non-zero dpendence of brain size on group size. Adding the phylogenentic infromation reduces evidence of a relationship between brain size and group size.







