---
title: "biol792ch3"
author: "morgan hrivnak"
date: "2024-09-10"
output: html_document
---
3e1. 
```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 6 , size=9 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )

sum( samples < 0.2 )
sum( samples < 0.2 ) / 1e4
```
3e2.
```{r}
sum( samples > 0.8) /1e4
```
3e3. 
```{r}
sum( samples > 0.2 & samples < 0.8 ) / 1e4
```
3e4.
```{r}
quantile( samples , 0.2 )
```
3e5.
```{r}
quantile( samples , 0.8 )
```
3e6.
```{r}
HPDI( samples , prob=0.66 )
```
3e7. 
```{r}
PI( samples , prob=0.66 )
```
3m1.
```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

plot( posterior ~ p_grid , type="l" )

```
3m2.
```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- rep( 1 , 1000 )
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)

samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
HPDI( samples , prob=0.9 )
```
3m3. 
```{r}
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
w <- rbinom( 1e4 , size=15 , prob=samples )
sum(( w==8 ) / 1e4)
```
3m4.
```{r}
w <- rbinom( 1e4 , size=9 , prob=samples )
sum( w==6 ) / 1e4
```
3m5.
```{r}
p_grid <- seq( from=0 , to=1 , length.out=1000 )
prior <- ifelse( p_grid < 0.5 , 0 , 1 )
likelihood <- dbinom( 8 , size=15 , prob=p_grid )
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
plot( posterior ~ p_grid , type="l" )

```
```{r}
HPDI( samples , prob=0.9 )
```
The prior is narrower that allows us to understand the values of below 0.5. Narrow priors help avoid bias into analysis. The above shows a slightly narrower prior that is more precise than what was previously calculated.

3m6.
how do we determine this with no prior sample size?
where code? how code?

there is an algebraic approach to find this that makes sense but that isn't how you would do it in r.

*increase sample size over and over again to get the confidence interval as close as possible

```{r}
p_grid <= seq(from=0 , to=1 , length.out=100)
prior <- rep(1,100)
likelihood <-dbinom(1600, size=3000, prob=p_grid)
posterior <- likelihood * prior
posterior <- posterior / sum(posterior)
set.seed(100)
samples <- sample( p_grid , prob=posterior , size=1e4 , replace=TRUE )
```






