---
title: "biol792ch4hw"
author: "morgan hrivnak"
date: "2024-09-18"
output: html_document
---
4E1. 

yi ∼ Normal(µ,σ) is likelhihood 
µ ∼Normal(0,10) is prior for paramenter µ
σ ∼Exponential(1) is the prior for parameter σ

4E2. 

2 normal parameters µ and σ

4E3.

Pr(µ,σ|y)= ((Pr(y|µ,σ)Pr(y))/((Pr(y|µ,σ)(Pr(µ)Pr(σ))

4E4. 

The linear model is represented in the second line. The other lines are potential parameters for the linear model.

4E5. 

There are 3 parameters represented by α, β, and σ.

4M1. 
```{r}
library(rethinking)
mu_prior <- rnorm( 1e4 , 0 , 10 )
sigma_prior <- runif( 1e4 , 0, 10 )
h_sim <- rnorm( 1e4 , mu_prior , sigma_prior )
dens( h_sim )
```
4M2.
```{r}

f <- alist(
y ~ dnorm( mu , sigma ),
mu ~ dnorm( 0 , 10 ),
sigma ~ dunif( 0 , 10 )
)

```
4M3. 
 y ~ dnorm( mu , sigma )       ->    yi ∼ Normal(µ, σ)
 mu <- a + b*x                 ->    µ = α + βxi
 a ~ dnorm( 0 , 10 )           ->    α ∼ Normal(0, 50)
 b ~ dunif( 0 , 1 )            ->    β ∼ Normal(0, 10)
 sigma ~ dexp( 1 )             ->    σ ∼ Uniform(0, 50) 
 
4M4. 

hi ∼ Normal(µ, σ)
µ = α + βyi
α ∼ Normal(0, 100)
β ∼ Uniform(0, 10)
σ ∼ Uniform(0, 50)

4M5. 

Not necceasirly, it wouldn't need to change priors generally. Generally speaking people do not get exponentially taller, people will stop growing depending on other factors so we can assume some individuals will grow over the three years other will not. This might look differently depending what kind of students are being measured, if they are elementary or middle school aged students they will grow more than students in high school and college who may be done growing.

4M6. 

The sigma may need to changed to be a range from (0,50) to (0.64) to encompass the entire range of heights of possible students growth.
